Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis

### 2021-03-23 manuscript/draft_rsfmri.Rmd
changed reference to latex compatible

---
title: "Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Evelyn Medawar [1]
- Kristin Prehn [4][5]
- Jürgen Ordemann [8][9]
- Agnes Floel [4][5]
- A. Veronica Witte [3][1][2]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Clinic of Cognitive Neurology, University Clinic Leipzig \newline 4 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 5 Department of Psychology, Medical School Hamburg, Hamburg \newline 6 Department of Neurology, University of Greifswald, Greifswald \newline 7 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin
output:
  bookdown::pdf_book:
    extra_dependencies: "subfig"
    fig_caption: yes
    number_sections: yes
    toc: yes
bibliography: "bibliography.bib"
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{setspace}
- \onehalfspacing
indent: true
---

@@

library(sjPlot) #only works for HTML output
library(texreg) #works for Rmarkdown
library(knitr) # version 1.30

@@

Obesity is a worldwide health issue, entailing huge personal and societal costs. Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life [@Albanese_2017]. Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI ($> 35$ kg/m$^2$). Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status. RYBG is the most invasive, yet most effective surgical procedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].\
Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla_2017; @Brutman_2019]. Precise mechanisms how bariatric surgery leads to altered appetitive signaling are yet to be elucidated. One option to adress these questions on brain-behavior relationships is to use resting-state functional magnetic resonance imaging (rsfMRI), a technique capturing the dynamic organization of the brain. Functional connectivity (FC) networks, i.e. brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009]. The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.\
The RN comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011]. These brain regions have been suggested to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018]. Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020]. RsfMRI studies also showed increased local FC of reward network regions, i.e. NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015]. In bariatric surgery patients, connectivity  within the RN (e.g. putamen, OFC) might be normalized by the surgery, however the evidence is limited due to a lack of longitudinal obese control groups  [@Duan_2020; @Wiemerslage_2016]. Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling , e.g. changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018].\
The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle_2015]. Higher BMI and obesity have been associated with a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e. salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018]. After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].\
In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive. Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].\
Further, while higher BMI has been consistently associated with more head motion during rsfMRI [@Beyer_2017;@Hodgson_2016], previous studies in bariatric surgery patients have not taken this important confounder of FC into account. In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020]. Thus, we aimed to rigorously control for motion-related variance in our analyses.\

@@

<!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité University Medicine Berlin, Germany. For more details see [@Prehn_2020]. We used all data acquired until April 2019.\

@@

Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin. Inclusion criteria were, in accordance with German guidelines for bariatric surgery, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g. type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013]. Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score $< 34$).\

@@

Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` females) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` females), who waited for their health insurance's approval to undergo surgery. Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016]. 

@@

Participants arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast. They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast. MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020).

@@

MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = $1 \times 1 \times 1$ mm$^3$, 192 sagital slices).
<!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms. The image matrix was $64 \times 64$ with an in-plane resolution of 3 mm $\times$ 3 mm and 34 slices with a slice thickness of 4 mm. 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes. Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneities.
Participants were instructed to close their eyes but to remain awake during scanning.

@@

### Minimal preprocessing
Imaging data analysis was conducted using AFNI 19.1.05, ANTS 2.3.1, FSL 6.0.1 and FreeSurfer 6.0.0p1, wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15 which can be found on https://github.com/fBeyer89/ADI_preproc/.
T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross]. Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long]. Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.  
The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 $1 \times 1 \times 1$ mm template using ANTS [@avants2009ants]. 
Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step. 
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).

### Denoising pipelines
In the pre-registration, we specified two denoising pipelines (AROMA+CC including ICA-AROMA + CompCor and AROMA+CC+GSR including ICA-AROMA + CompCor + global signal regression (GSR)) [@Parkes_2018; @Ciric_2017, for details see SI].

@@

**Anatomical quality control**
FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019. We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.\
<!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
**RsfMRI quality control**
RsfMRI quality control was performed according to the protocol by @Ciric_2018 (see SI for more details). 
Head motion was quantified using mean framewise displacement (mFD) according to @Power_2012, and log-transformed for further analysis (logmFD).
As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].

@@

Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al. -@spm12). The marginal model implemented in the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size. We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to $3 \times 3 \times 3$ mm$^3$ and thresholded at 0.5 GM probability) for all analyses. Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].

<!-- MODEL SETUP --> 
### Confirmatory analyses
**Whole-brain analysis**
We tested the pre-registered hypothesis of a time-by-group interaction:  `FC = group + time + group * time`.\
As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both time points (model CA2) and age, sex, average logmFD and baseline BMI (model CA3). 
Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values. This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.\
**Aggregated FC analysis**
We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007]. We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline. First, we investigated the time-by-group interaction for baseline and followup time points only. We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3). We performed model comparison between R1 and R0 models, where `R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))`, and `R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))`. As specified in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.

### Exploratory analysis (pre-registered)
**Whole-brain analysis**
As described in the pre-registration, we performed an exploratory analysis of the relationship of average and change in BMI in a separate group of models. As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI. This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC. We first estimated their effects in a model adjusting for baseline age and sex (model EA1) and then additionally controlling for logmFD (model EA2). As we previously reported correlated change in BMI and head motion in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (model EA3). Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.\
**Aggregated FC analysis**
We also performed an exploratory analysis with average BMI and change in BMI as predictors of the aggregated FC from AROMA+CC+GSR denoised data of both networks. We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (model EA1), logmFD (model EA2) and average and change in logmFD (model EA3).

### Exploratory analysis (not pre-registered)
**Whole-brain analysis**
**Group-by-time interaction in three time points**
We explored our whole-brain group-by-time point interaction models for the data of all three time points. In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017]. The resulting factorial design contained one regressor for each time point per group (see SI for depiction of design matrix). This analysis had not been pre-registered. Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points. We investigated two models adjusting for age and sex (model EA4) and age, sex and logmFD (model EA5).\
**Average and change in logmFD**
Further, we tested the association of head motion and FC in another exploratory model (model EA6): `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates. With this model, we aimed for a better understanding of the unique contribution of average and longitudinal change in logmFD measures.

@@

**Whole-brain analysis**
<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference. Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual. Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$), and 1000 bootstraps due to required computation time (instead of 5000). Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$. The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
**Aggregated FC analysis**
The interaction effect of group and time point in the models of aggregated FC was considered significant if the model comparison between R1 and R0 models using the anova command showed $p < 0.05$. In all exploratory models, we considered all coefficients with $p < 0.05$ as significant.

### Functional decoding
In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the NeuroSynth (https://www.neurosynth.org/) database [@Yarkoni_2011]. We uploaded the contrast images for change BMI and average logmFD on Neurovault (https://neurovault.org/collections/9426/), and applied the decoding classifier. This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps. We reported the three top terms for both contrasts.

@@

In this study, we could not confirm our pre-registered hypotheses. Based on previous studies in bariatric surgery patients, we expected within-DMN FC to increase, and DMN FC to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013]. Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019]. These studies, notably, did not include adequate longitudinal obese control groups, making false-positive findings possible. We thus conclude that surgery-induced heavy weight loss does not strongly affect DMN and RN FC based on the current results.\

@@

One the other hand, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning. Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi_2018]. Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging. Importantly, a global decrease in connectivity might go along with selectively increased connectivity, i.e. a shift from integration towards segregation and more specialized processing. However, this interpretation is highly speculative, and more studies are needed to corroborate these considerations.\
Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity, challenging definite conclusions on the relationship between weight loss and FC changes. Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.\

>>>>

---
title: "Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis"
author:
- Hannah Sophie Heinrichs* 
- Frauke Beyer* 
- Evelyn Medawar
- Kristin Prehn
- Jürgen Ordemann
- Agnes Floel
- A. Veronica Witte
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline 2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Clinic of Cognitive Neurology, University Clinic Leipzig \newline 4 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 5 Department of Psychology, Medical School Hamburg, Hamburg \newline 6 Department of Neurology, University of Greifswald, Greifswald \newline 7 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin

bookdocumentclass: article
output:
  bookdown::pdf_book:
    extra_dependencies: "subfig"
    fig_caption: yes
    number_sections: yes
    toc: yes
    #toc_depth: 4
    keep_tex: yes
    #citation_package: none
    pandoc_args: [ "--csl","apa.csl" ]
bibliography: "bibliography.bib"
link-citations: true

editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{setspace}
- \onehalfspacing
#- \usepackage{bibtex}
#indent: true
---

@@

library(sjPlot) #only works for HTML output
library(texreg) #works for Rmarkdown
library(knitr) # version 1.30
library(distill)

@@

Obesity is a worldwide health issue, entailing huge personal and societal costs. Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life [@Albanese2017]. Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI ($> 35$ kg/m$^2$). Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status. RYBG is the most invasive, yet most effective surgical procedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].\
Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla2017; @Brutman2019]. Precise mechanisms how bariatric surgery leads to altered appetitive signaling are yet to be elucidated. One option to address these questions on brain-behavior relationships is to use resting-state functional magnetic resonance imaging (rsfMRI), a technique capturing the dynamic organization of the brain. Functional connectivity (FC) networks, i.e. brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith2009]. The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.\
The RN comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty2004;@Liu2011]. These brain regions have been suggested to guide food valuation processes and decision-making in humans [@Bartra2013; @Hare2011; @Hutcherson2012; @Schmidt2018]. Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto2018; @Garcia2014; @Meng2020; @Stoeckel2009; though see @Morys2020]. RsfMRI studies also showed increased local FC of reward network regions, i.e. NAcc, vmPFC, putamen, insula [@Contreras2017; @Coveleskie2015;@Hogenkamp2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips2014;  @Wijngaarden2015]. In bariatric surgery patients, connectivity  within the RN (e.g. putamen, OFC) might be normalized by the surgery, however the evidence is limited due to a lack of longitudinal obese control groups  [@Duan2020; @Wiemerslage2016]. Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling, e.g. changes in ghrelin levels, via hypothalamic-striatal projections [@Li2019; @Karra2013; but @Zoon2018].\
The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle2015]. Higher BMI and obesity have been associated with a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e. salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer2017; @Kullmann2011; @Chao2018; @Wijngaarden2015; @Borowitz2020; @Ding2020; @Doucet2017; @Sadler2018]. After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li2018; @Olivo2017; @Frank2013].\
In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive. Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George2016].\
Further, while higher BMI has been consistently associated with more head motion during rsfMRI [@Beyer2017;@Hodgson2016], previous studies in bariatric surgery patients have not taken this important confounder of FC into account. In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer2020]. Thus, we aimed to rigorously control for motion-related variance in our analyses.\

@@

<!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité University Medicine Berlin, Germany. For more details see [@Prehn2020]. We used all data acquired until April 2019.\

@@

Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin. Inclusion criteria were, in accordance with German guidelines for bariatric surgery, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g. type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick2013]. Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score $< 34$).\

@@

Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` females) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` females), who waited for their health insurance's approval to undergo surgery. Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski2016]. 

@@

Participants arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast. They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast. MRI scanning was done after performing a psychological test battery (for details see @Prehn2020).

@@

MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = $1 \times 1 \times 1$ mm$^3$, 192 sagital slices).
<!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms. The image matrix was $64 \times 64$ with an in-plane resolution of 3 mm $\times$ 3 mm and 34 slices with a slice thickness of 4 mm. 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes. Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneities.
Participants were instructed to close their eyes but to remain awake during scanning.

@@

### Minimal preprocessing
Imaging data analysis was conducted using AFNI 19.1.05, ANTS 2.3.1, FSL 6.0.1 and FreeSurfer 6.0.0p1, wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15 which can be found on https://github.com/fBeyer89/ADI_preproc/.
T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross]. Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fslong]. Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.  
The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 $1 \times 1 \times 1$ mm template using ANTS [@avants2009ants]. 
Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step. 
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).

### Denoising pipelines
In the pre-registration, we specified two denoising pipelines (AROMA+CC including ICA-AROMA + CompCor and AROMA+CC+GSR including ICA-AROMA + CompCor + global signal regression (GSR)) [@Parkes2018; @Ciric2017, for details see SI].

@@

**Anatomical quality control**
FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk2019. We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.\
<!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
**RsfMRI quality control**
RsfMRI quality control was performed according to the protocol by @Ciric2018 (see SI for more details). 
Head motion was quantified using mean framewise displacement (mFD) according to @Power2012, and log-transformed for further analysis (logmFD).
As pre-registered, we did not exclude anybody based on high average head motion [@Beyer2020].

@@

Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al. -@spm12). The marginal model implemented in the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size. We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to $3 \times 3 \times 3$ mm$^3$ and thresholded at 0.5 GM probability) for all analyses. Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].

<!-- MODEL SETUP --> 
### Confirmatory analyses
**Whole-brain analysis**
We tested the pre-registered hypothesis of a time-by-group interaction:  `FC = group + time + group * time`.\
As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both time points (model CA2) and age, sex, average logmFD and baseline BMI (model CA3). 
Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values. This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.\
**Aggregated FC analysis**
We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates2007]. We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline. First, we investigated the time-by-group interaction for baseline and followup time points only. We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3). We performed model comparison between R1 and R0 models, where `R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))`, and `R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))`. As specified in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.

### Exploratory analysis (pre-registered)
**Whole-brain analysis**
As described in the pre-registration, we performed an exploratory analysis of the relationship of average and change in BMI in a separate group of models. As proposed by @Guillaume2014, we calculated the between- and within subject centered values of BMI. This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC. We first estimated their effects in a model adjusting for baseline age and sex (model EA1) and then additionally controlling for logmFD (model EA2). As we previously reported correlated change in BMI and head motion in this sample [@Beyer2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (model EA3). Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.\
**Aggregated FC analysis**
We also performed an exploratory analysis with average BMI and change in BMI as predictors of the aggregated FC from AROMA+CC+GSR denoised data of both networks. We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (model EA1), logmFD (model EA2) and average and change in logmFD (model EA3).

### Exploratory analysis (not pre-registered)
**Whole-brain analysis**
**Group-by-time interaction in three time points**
We explored our whole-brain group-by-time point interaction models for the data of all three time points. In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo2017]. The resulting factorial design contained one regressor for each time point per group (see SI for depiction of design matrix). This analysis had not been pre-registered. Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points. We investigated two models adjusting for age and sex (model EA4) and age, sex and logmFD (model EA5).\
**Average and change in logmFD**
Further, we tested the association of head motion and FC in another exploratory model (model EA6): `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates. With this model, we aimed for a better understanding of the unique contribution of average and longitudinal change in logmFD measures.

@@

**Whole-brain analysis**
<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference. Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual. Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$), and 1000 bootstraps due to required computation time (instead of 5000). Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$. The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
**Aggregated FC analysis**
The interaction effect of group and time point in the models of aggregated FC was considered significant if the model comparison between R1 and R0 models using the anova command showed $p < 0.05$. In all exploratory models, we considered all coefficients with $p < 0.05$ as significant.

### Functional decoding
In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the NeuroSynth (https://www.neurosynth.org/) database [@Yarkoni2011]. We uploaded the contrast images for change BMI and average logmFD on Neurovault (https://neurovault.org/collections/9426/), and applied the decoding classifier. This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps. We reported the three top terms for both contrasts.

@@

In this study, we could not confirm our pre-registered hypotheses. Based on previous studies in bariatric surgery patients, we expected within-DMN FC to increase, and DMN FC to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li2018; @McFadden2013; @Frank2013]. Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat [@Ochner2012_wanting; @Scholtz2014; @Cerit2019]. These studies, notably, did not include adequate longitudinal obese control groups, making false-positive findings possible. We thus conclude that surgery-induced heavy weight loss does not strongly affect DMN and RN FC based on the current results.\

@@

One the other hand, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning. Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi2018]. Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging. Importantly, a global decrease in connectivity might go along with selectively increased connectivity, i.e. a shift from integration towards segregation and more specialized processing. However, this interpretation is highly speculative, and more studies are needed to corroborate these considerations.\
Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity, challenging definite conclusions on the relationship between weight loss and FC changes. Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer2020], and therefore conducted careful analyses of the impact of head motion on our results.\
